---
title: "Figures3CEF"
author: "Joe Stock"
date: "December 8, 2017"
output: html_document
---

# Set the working directory for the files we'll be working with.

```{r}
setwd("C:/Users/Stock/Desktop/BCB-546X/BCB546X-Fall2017/Final_Group_Project/")
```

# Differentially expressed genes in cluster 11 expressed in all mutants
# Read in the supplementary files for the first graph.

# Load libraries

```{r}
library(readxl)
library(readr)
library(gtools)
library(reshape2)
library(scales)
library(plyr)
library(ggplot2)
```


# Load the necessary files

```{r}
Supp_Table5 <- read_delim("Supp_Table5.txt", 
                          "\t", escape_double = FALSE, trim_ws = TRUE)
                          
Supp_Table7 <- read_excel("Supp_Table7.xlsx")
Supp_Table3 <- read_excel("Supp_Table3.xlsx")
Supp_Table6 <- read_excel("Supp_Table6.xlsx")
```

# Because we are only interested in the genes that are signifcantly differentially expressed in each mutant,
# we can remove any from Table 7 that do not have fold changes values for each mutant.

```{r}
Supp_Table7 <- Supp_Table7[complete.cases(Supp_Table7[ , 4:7]),]
```

# Remove the first duplicated row of the data frame

```{r}
Supp_Table5 <- Supp_Table5[-1,]
```

# Now we can pull the values from Supplementary Table 5 to calculate the fold change differences and their
# log scores to develop the heat map.
# What we are trying to accomplish with this is reducing whole Table 5 data frame down, to only genes 
# that are differentially expressed in group 11 for each mutant.

```{r}
Supp_Table <- subset(Supp_Table5, Supp_Table5$gene.id %in% Supp_Table7$maize.gene.id)
head(Supp_Table)
```

# Dim() will define whether we have achieved our goal.

```{r}
dim(Supp_Table)
dim(Supp_Table7)
```

# As we can see each data frame is now 28 obs long, which coincides with the same genes in each

# I do not need all of the variables to calculate fold changes for the mutants vs wild type, so 
# I will remove those that I do not need. 

```{r}
Supp_Table_clean <- Supp_Table[,-c(2:7,12:15)]
```

# Now I will have to calculate the fold changes and convert to log scale. I need to install the gtools
# package to be able to use the fold change package.

```{r}
Supp_Table_clean$ra1_fc <- foldchange(as.numeric(Supp_Table_clean$ear.ra1.1mm),as.numeric(Supp_Table_clean$ear.wt.1mm))
Supp_Table_clean$ra1_lnfc <- log(abs(Supp_Table_clean$ra1_fc))
Supp_Table_clean$ra1_lnfc <- ifelse(Supp_Table_clean$ra1_fc < 0, -1 * Supp_Table_clean$ra1_lnfc, Supp_Table_clean$ra1_lnfc) 

Supp_Table_clean$ra2_fc <- foldchange(as.numeric(Supp_Table_clean$ear.ra2.1mm),as.numeric(Supp_Table_clean$ear.wt.1mm))
Supp_Table_clean$ra2_lnfc <- log(abs(Supp_Table_clean$ra2_fc))
Supp_Table_clean$ra2_lnfc <- ifelse(Supp_Table_clean$ra2_fc < 0, -1 * Supp_Table_clean$ra2_lnfc, Supp_Table_clean$ra2_lnfc)

Supp_Table_clean$ra3_fc <- foldchange(as.numeric(Supp_Table_clean$ear.ra3.1mm),as.numeric(Supp_Table_clean$ear.wt.1mm))
Supp_Table_clean$ra3_lnfc <- log(abs(Supp_Table_clean$ra3_fc))
Supp_Table_clean$ra3_lnfc <- ifelse(Supp_Table_clean$ra3_fc < 0, -1 * Supp_Table_clean$ra3_lnfc, Supp_Table_clean$ra3_lnfc)

Supp_Table_Final <- Supp_Table_clean[,-c(2:6,8,10)]

Supp_Table_Final <- Supp_Table_Final[order(-Supp_Table_Final$ra3_lnfc),]
```

# Because I am getting return errors for the value of -Inf in my melt file, it messes up the 
# transformation later, therefore I manually deleted the measure

```{r}
Supp_Table_Final <- Supp_Table_Final[-28,]


Supp_Table_Final.m <- melt(Supp_Table_Final)


Supp_Table_Final.m <- ddply(Supp_Table_Final.m, .(variable), transform, rescale = rescale(value))

dat.temp <- Supp_Table_Final.m %>%
  filter(variable %in% c("ra3_lnfc")) %>%
  arrange(rescale)

Supp_Table_Final.m$gene.id <- factor(Supp_Table_Final.m$gene.id, levels=as.character(dat.temp$gene.id))


(p <- 
    ggplot(Supp_Table_Final.m, aes(variable, gene.id)) + 
    geom_tile(aes(fill=value), colour = "black") + 
    scale_fill_gradient2(low = "red", mid = "black", high = "green")
  )
```

# Figure 3E - Differentially expressed genes in cluster 8 expressed in all mutants

```{r}
Supp_Table3 <- Supp_Table3[complete.cases(Supp_Table3[ , 6:8]),]  
```

# Pull cluster 8 observations from table 6 and identify complete cases across all mutants

```{r}
Clust8 <- Supp_Table6[Supp_Table6$clustera == 8,]
Clust8$DE.ra1.1mm[Clust8$DE.ra1.1mm == "no"] <- NA
Clust8$DE.ra2.1mm[Clust8$DE.ra2.1mm == "no"] <- NA
Clust8$DE.ra3.1mm[Clust8$DE.ra3.1mm == "no"] <- NA

Clust8 <- Clust8[complete.cases(Clust8[ , 13:15]),] 
```

# Organizing and cleaning

# This methods pulls DE from Table 3, however when used the values do not make sense.
# Supp_Table3$id <- substr(Supp_Table3$id,1,13)
# Supp_Table3$id <- toupper(Supp_Table3$id)
# Clust8_Sub <- subset(Supp_Table3, Supp_Table3$id %in% Supp_Table6$maize.gene.id)
# Clust8_Sub <- Clust8_Sub[,-9]
# Clust8_Sub <- Clust8_Sub[,-c(1,2,4,5)]
# Clust8_Sub.m <- melt(Clust8_Sub)
# Clust8_Sub.m <- ddply(Clust8_Sub.m, .(variable), transform, rescale = rescale(value))

```{r}
Cluster8 <- subset(Supp_Table5, Supp_Table5$gene.id %in% Clust8$maize.gene.id)

Cluster8$ra1_fc <- foldchange(as.numeric(Cluster8$ear.ra1.1mm),as.numeric(Cluster8$ear.wt.1mm))
Cluster8$ra1_lnfc <- log(abs(Cluster8$ra1_fc))
Cluster8$ra1_lnfc <- ifelse(Cluster8$ra1_fc < 0, -1 * Cluster8$ra1_lnfc, Cluster8$ra1_lnfc) 

Cluster8$ra2_fc <- foldchange(as.numeric(Cluster8$ear.ra2.1mm),as.numeric(Cluster8$ear.wt.1mm))
Cluster8$ra2_lnfc <- log(abs(Cluster8$ra2_fc))
Cluster8$ra2_lnfc <- ifelse(Cluster8$ra2_fc < 0, -1 * Cluster8$ra2_lnfc, Cluster8$ra2_lnfc)

Cluster8$ra3_fc <- foldchange(as.numeric(Cluster8$ear.ra3.1mm),as.numeric(Cluster8$ear.wt.1mm))
Cluster8$ra3_lnfc <- log(abs(Cluster8$ra3_fc))
Cluster8$ra3_lnfc <- ifelse(Cluster8$ra3_fc < 0, -1 * Cluster8$ra3_lnfc, Cluster8$ra3_lnfc)

Cluster8 <- Cluster8[-5,]
Cluster8 <- Cluster8[,-c(2:16,18,20)]

Cluster8.m <- melt(Cluster8)
Cluster8.m <- ddply(Cluster8.m, .(variable), transform, rescale = rescale(value))

(p <- 
    ggplot(Cluster8.m, aes(variable, gene.id)) + 
    geom_tile(aes(fill=value), colour = "black") + 
    scale_fill_gradient2(low = "red", mid = "black", high = "green")
)
```

# Genes DE in ra1 and expression vs wild type

```{r}
Clust8DEra1 <- Supp_Table6[Supp_Table6$clustera == 8,]
Clust8DEra1$DE.ra1.1mm[Clust8DEra1$DE.ra1.1mm == "no"] <- NA

Clust8DEra1 <- Clust8DEra1[complete.cases(Clust8DEra1[ , 13]),] 

Cluster8DEra1 <- subset(Supp_Table5, Supp_Table5$gene.id %in% Clust8DEra1$maize.gene.id)

Cluster8DEra1$ra1_fc <- foldchange(as.numeric(Cluster8DEra1$ear.ra1.1mm),as.numeric(Cluster8DEra1$ear.wt.1mm))
Cluster8DEra1$ra1_lnfc <- log(abs(Cluster8DEra1$ra1_fc))
Cluster8DEra1$ra1_lnfc <- ifelse(Cluster8DEra1$ra1_fc < 0, -1 * Cluster8DEra1$ra1_lnfc, Cluster8DEra1$ra1_lnfc) 

Cluster8DEra1$ra2_fc <- foldchange(as.numeric(Cluster8DEra1$ear.ra2.1mm),as.numeric(Cluster8DEra1$ear.wt.1mm))
Cluster8DEra1$ra2_lnfc <- log(abs(Cluster8DEra1$ra2_fc))
Cluster8DEra1$ra2_lnfc <- ifelse(Cluster8DEra1$ra2_fc < 0, -1 * Cluster8DEra1$ra2_lnfc, Cluster8DEra1$ra2_lnfc)

Cluster8DEra1$ra3_fc <- foldchange(as.numeric(Cluster8DEra1$ear.ra3.1mm),as.numeric(Cluster8DEra1$ear.wt.1mm))
Cluster8DEra1$ra3_lnfc <- log(abs(Cluster8DEra1$ra3_fc))
Cluster8DEra1$ra3_lnfc <- ifelse(Cluster8DEra1$ra3_fc < 0, -1 * Cluster8DEra1$ra3_lnfc, Cluster8DEra1$ra3_lnfc)

Cluster8DEra1 <- Cluster8DEra1[,-c(2:16,18,20)]

Cluster8DEra1.m <- melt(Cluster8DEra1)
Cluster8DEra1.m <- ddply(Cluster8DEra1.m, .(variable), transform, rescale = rescale(value))

(p <- 
    ggplot(Cluster8DEra1.m, aes(variable, gene.id)) + 
    geom_tile(aes(fill=value), colour = "black") + 
    scale_fill_gradient2(low = "red", mid = "black", high = "green")
)
```


