---
title: "Figure_2_D_E"
author: "Anita Corredor"
date: "12/8/2017"
output: html_document
---
Loading Packages:
```{r }
library(readxl)
library(ggplot2)
library(reshape)
library(reshape2)
library(plyr)
library(scales)
```

#Figure 2 D
Data is located in test5.xlsx which is modified from supp_gr.166397.113_Supp_Table3.xlsx

##prepare the data for heatmap
```{r}
test5 <- data.frame(test5)
test5.m<- melt(test5)
test5.m$gen2<-factor(test5.m$gen, levels=rev(levels(factor(test5.m$gen))))
```

##generating the plot with ggplot2 package
```{r}
(s <- ggplot(test5.m, aes(variable, gen2)) + geom_tile(aes(fill = value), 
              colour = "white") + 
    scale_fill_gradient2(low = "red", mid = "white", high = "blue")   )
s+labs(x="",y="") 
```



#Figures 2 E-1
We create a Data set with the names of the families that are going to be used on the figures TCP.xlsx and MADS.xlsx which are extracted from Supp_Table3.xlsx

##prepare the data for heatmap
```{r}
toupper(TCP$id) 
TCP$id <- toupper(TCP$id)
```

##subsetting the data 
We used the information of the values of ln fold change from Supplemental_Table_2
```{r}
Supp_Table2_TCP <- subset(Supplemental_Table_2, Supplemental_Table_2$gene_id %in% TCP$id)
Supp_Table2_TCP <- Supp_Table2_TCP[,-c(2,4,6,8,10:15)]
remove <- NA
Supp_Table2_TCP <- Supp_Table2_TCP[!Supp_Table2_TCP$size %in% remove,]
remove<-"tas"
Supp_Table2_TCP <- Supp_Table2_TCP[!Supp_Table2_TCP$q2 %in% remove,]
```

##arranging the levels in the same order as on the original figure
```{r}
Supp_Table2_TCP$gene_id<-factor(Supp_Table2_TCP$gene_id,levels=c("GRMZM2G178603",
                                                                 "GRMZM2G465091",
                                                                 "GRMZM2G078077",
                                                                 "GRMZM2G089361",
                                                                 "AC205574.3_FGT006",
                                                                 "AC199782.5_FGT003",
                                                                 "GRMZM2G092214",
                                                                 "GRMZM2G089638",
                                                                 "GRMZM2G093895",
                                                                 "GRMZM2G015037",
                                                                 "GRMZM2G096610",
                                                                 "GRMZM2G064628",
                                                                 "GRMZM2G330719"
                                                                 ))
```


##arranging the siginificance sign on the Subset Data 
```{r}
Supp_Table2_TCP$significant[Supp_Table2_TCP$significant == "no"] <- ""
Supp_Table2_TCP$significant[Supp_Table2_TCP$significant == "yes"] <- "*"
```

##arranging the mutants ra1, ra2 and ra3 with the size expression on mm
```{r}
Supp_Table2_TCP$sizeq2 <- paste(Supp_Table2_TCP$q2, Supp_Table2_TCP$size)

```

##plotting the heat map for the TCP family of genes
```{r}
(t <- 
    ggplot(Supp_Table2_TCP, aes(sizeq2, gene_id)) + 
    geom_tile(aes(fill=ln_fc), colour = "white") + 
    scale_fill_gradient2(low = "red", mid = "white", high = "blue") +
    geom_text(aes(label = significant), colour = "black", fontface = "bold")
)
```

##modifing the labels for final presentation
```{r}

t+labs(x="* DE p < 0.05;FDR 10%",y="DE TCP TFs")
```


#Figures 2 E-2

##prepare the data for heatmap                                                                 
```{r}                                                                 
toupper(MADS$id) 
MADS$id <- toupper(MADS$id)
```

##subsetting the data 
We used the information of the values of ln fold change from Supplemental_Table_2
```{r}
Supp_Table2_MADS <- subset(Supplemental_Table_2, Supplemental_Table_2$gene_id %in% MADS$id)
Supp_Table2_MADS <- Supp_Table2_MADS[,-c(2,4,6,8,10:15)]
remove <- NA
Supp_Table2_MADS <- Supp_Table2_MADS[!Supp_Table2_MADS$size %in% remove,]
remove<-"2mm"
Supp_Table2_MADS<-Supp_Table2_MADS[!Supp_Table2_MADS$size %in% remove,]
```

##arranging the levels in the same order as on the original figure
```{r}
Supp_Table2_MADS$gene_id<-factor(Supp_Table2_MADS$gene_id, levels=c("GRMZM2G117961",
                                                                    "GRMZM2G370777",
                                                                    "GRMZM2G071620",
                                                                    "GRMZM2G059102",
                                                                    "GRMZM2G046885",
                                                                    "GRMZM2G033093",
                                                                    "GRMZM5G805387",
                                                                    "GRMZM2G026223",
                                                                    "AC234185.1_FGT006",
                                                                    "GRMZM2G105387",
                                                                    "GRMZM5G891280",
                                                                    "GRMZM2G005155"
                                                                    ))
```

##arranging the siginificance sign on the Subset Data 
```{r}
Supp_Table2_MADS$significant[Supp_Table2_MADS$significant == "no"] <- ""
Supp_Table2_MADS$significant[Supp_Table2_MADS$significant == "yes"] <- "*"
```

##plotting the heat map for the MADS family of genes
```{r}
(u <- 
    ggplot(Supp_Table2_MADS, aes(q2, gene_id)) + 
    geom_tile(aes(fill=ln_fc), colour = "white") + 
    scale_fill_gradient2(low = "red", mid = "white", high = "blue") +
    geom_text(aes(label = significant), colour = "black", fontface = "bold")
)
```

##modifing the labels for final presentation
```{r}
u+labs(x="",y="DE MADS-box TFs")
```






