---
title: "Figure3_A to C&G"
author: "Yan Zhou"
date: "12/7/2017"
output: html_document
---
Loading Packages:
```{r }
library(reshape2)
library(ggplot2)
library(gridExtra)
library(grid)
library(cowplot)
library(Rmisc)
```

## Figure 3 A

Data is located in expression_table_s6.csv which is modified from S_table6.xlx

```{r}
kc<-read.csv("expression_table_s6.csv",header = T)
p<-kc$clustera
p<-as.data.frame(p)
names(p)<-"cluster"

```
#generating the FPKM matrix by removing the cluster column 
#to calculate the cluster, it is first make the z score of each gene' relatively expression leve in tissues.
#so we need to cauculate the z score row by row
#because the scale fuction can't calculate column by column, we need first tranform the the data frame 
#then transform the data back to original order
```{r}
kk<-kc[,-1]

kw<-NULL

kk1<-t(kk)
kw<-NULL
for (i in 1:ncol(kk1)) {
  km<-NULL
  kn<-kk1[,i]
  k<-scale(kn)
  #get the z score
  km<- (k-mean(k))/sd(k)
  kw<-cbind(kw,km)
}
#transform the data back to original order
kw2<-t(kw)
kw2<-as.data.frame(kw2)
#adding the cluster information back
kw2$clustera<-p$cluster
m<-table(kw2$clustera)
m<-as.data.frame(m)
kw2$clustera<-as.character(kw2$clustera)
#calculating the mean z score of each element
w<-aggregate(kw2,by=list(kw2$clustera), FUN=mean)
```

#prepare the data for heatmap
```{r}
names(w)<-names(kc)
w<-w[,c(1:9)]
#give the names of each clumn
names(w)<-c("clustera","1mm","2mm","IM/SPM","SM","FM","stg1","stg2","stg3")
df.melted <- melt(w, id.vars=c("clustera"))
names(df.melted)<-c("cls","ti","value")
df.melted$value=df.melted$value
```

#generating the heatmap of relative experession level of each cluster
```{r}
hm<-ggplot(data = df.melted, aes(x = factor(ti), y =factor(cls,levels=c("3", 
                                                                        "4", 
                                                                        "1",
                                                                        "17",
                                                                        "13",
                                                                        "7",
                                                                        "8",
                                                                        "15",
                                                                        "6",
                                                                        "9",
                                                                        "19",
                                                                        "12",
                                                                        "11",
                                                                        "16",
                                                                        "10",
                                                                        "5",
                                                                        "14",
                                                                        "18",
                                                                        "2",
                                                                        "20")), fill = value))+
  geom_tile()+
  scale_fill_distiller(name = "Legend title", palette = "Blues", direction = 1, na.value = "transparent") +
  scale_x_discrete(breaks = unique(df.melted$ti), labels = unique(df.melted$ti)) + theme_bw() +
  scale_x_discrete(position = "top")+
  xlab("")+
  ylab("")+
  theme(legend.position = "bottom", legend.direction = "horizontal",
        legend.title = element_text(size = 15), legend.key.size = unit(1,"cm"),
        legend.text = element_text(size = 7)) +
  guides(fill = guide_colorbar(title.position = "top", title.hjust = 0.5))

#modify the theme and axis
hm1<-hm+theme_minimal()+  
  theme(axis.text.x = element_text(angle = 60,hjust=0,face = "bold", size = 12))+
  theme(axis.text.y = element_text(face = "bold", size = 12))+
  theme(legend.position = "bottom", legend.direction = "horizontal",
        legend.title = element_text(size = 15), legend.key.size = unit(0.5,"cm"),
        legend.text = element_text(size = 7))+
        theme(legend.title=element_blank())
        guides(fill = guide_legend(title=""))
```

#addling a plot shows number of each cluster

```{r}

lab<-ggplot(data = m, aes(x = factor(k), y =factor(Var1,levels=c("3", 
                                                                 "4", 
                                                                 "1",
                                                                 "17",
                                                                 "13",
                                                                 "7",
                                                                 "8",
                                                                 "15",
                                                                 "6",
                                                                 "9",
                                                                 "19",
                                                                 "12",
                                                                 "11",
                                                                 "16",
                                                                 "10",
                                                                 "5",
                                                                 "14",
                                                                 "18",
                                                                 "2",
                                                                 "20"))))+
  geom_tile(fill="white") + 
  geom_text(aes(label = round(Freq, 1)))+
  xlab("")+
  ylab("")+theme_minimal()+
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())+
  theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())+
  theme(legend.position = "none")
```

# arrange the layout of combining plot to adjust the two figuress
```{r}
lay <- rbind(c(1,1,1,1,NA),
             c(1,1,1,1,NA),
             c(1,1,1,1,NA),
             c(1,1,1,1,NA),
             c(1,1,1,1,NA),
             c(1,1,1,1,2),
             c(1,1,1,1,2),
             c(1,1,1,1,2),
             c(1,1,1,1,2),
             c(1,1,1,1,2),
             c(1,1,1,1,2),
             c(1,1,1,1,2),
             c(1,1,1,1,2),
             c(1,1,1,1,2),
             c(1,1,1,1,2),
             c(1,1,1,1,2),
             c(1,1,1,1,2),
             c(1,1,1,1,2),
             c(1,1,1,1,2),
             c(1,1,1,1,2),
             c(1,1,1,1,2),
             c(1,1,1,1,2),
             c(1,1,1,1,2),
             c(1,1,1,1,2),
             c(1,1,1,1,2),
             c(1,1,1,1,2),
             c(1,1,1,1,2),
             c(1,1,1,1,2),
             c(1,1,1,1,NA),
             c(1,1,1,1,NA),
             c(1,1,1,1,NA),
             c(1,1,1,1,NA))

#combine the plot together

grid.arrange(hm1,lab, layout_matrix = lay)
```

## Firgure3 B&C

#generateing two plots of cluster 11 and 8
```{r}
c11<-subset(df.melted,cls=="11")
c8<-subset(df.melted,cls=="8")
```
#cluster 8
```{r}
c8.1<-ggplot(data = c8, aes(x = factor(ti), y =factor(cls), fill = value))+
  geom_tile(height=0.2)+
  scale_fill_distiller(name = "Legend title", palette = "Blues", direction = 1, na.value = "transparent") +
  scale_x_discrete(breaks = unique(df.melted$ti), labels = unique(df.melted$ti)) + theme_bw() +
  scale_x_discrete(position = "top")+
  xlab("")+
  ylab("")+
  guides(fill = guide_colorbar(title.position = "top", title.hjust = 0.5))+theme_minimal()+  
  theme(axis.text.x = element_text(angle = 60,hjust=0,face = "bold", size = 12))+
  theme(axis.text.y = element_blank())+
  theme(line = element_blank())+
  theme(legend.position = "none")
c8.2<-ggplot(data=c8,aes(x=factor(ti),y=value,group=1))+
  geom_point(colour="blue",size=2)+
  geom_line()+
  theme(axis.text.x = element_blank())+
  ylab("relative expression")+
  xlab("")

grid.arrange(c8.1, c8.2, nrow = 2)
```

#cluster 11
```{r}
c11.1<-ggplot(data = c11, aes(x = factor(ti), y =factor(cls), fill = value))+
  geom_tile(height=0.2)+
  scale_fill_distiller(name = "Legend title", palette = "Blues", direction = 1, na.value = "transparent") +
  scale_x_discrete(breaks = unique(df.melted$ti), labels = unique(df.melted$ti)) + theme_bw() +
  scale_x_discrete(position = "top")+
  xlab("")+
  ylab("")+
  guides(fill = guide_colorbar(title.position = "top", title.hjust = 0.5))+theme_minimal()+  
  theme(axis.text.x = element_text(angle = 60,hjust=0,face = "bold", size = 12))+
  theme(axis.text.y = element_blank())+
  theme(line = element_blank())+
  theme(legend.position = "none")
c11.2<-ggplot(data=c11,aes(x=factor(ti),y=value,group=1))+
  geom_point(colour="blue",size=2)+
  geom_line()+
  ylim(-1,3)+
  theme(axis.text.x = element_blank())+
  ylab("relative expression")+
  xlab("")
grid.arrange(c11.1, c11.2, nrow = 2)
```
##Figure3 G

#Following are plot for generating the experssion levele of different genes in cluster8. The fpkm value is a combination of sup tableS6 for tissue expression and S2 for mutants expression
```{r}
id<-read.csv("cluster8.csv",header = T)
#rename the cluster file
names(id)<-c("id","IM/SPM","SM","FM","stg1","stg2","stg3","wt_1mm", "ra1_1mm","ra2_1mm","ra3_1mm","wt_2mm","ra1_2mm","ra2_2mm","ra3_2mm")
#divide into 3 subgroups
id1<-subset(id,id=="rs2"|id=="CUC-like"|id=="TCP TF")
id2<-subset(id,id=="unknown grass-specific"|id=="ROXY")
id3<-subset(id,id=="wox9C"|id=="LOB TF")
#Assgin the subgroup ids so we can later use facet fuction of ggplot to generate 3 plots
m1<-melt(id1, id.vars=c("id"))
names(m1)<-c("id","ti","value")
m1$cluster=1
m2<-melt(id2, id.vars=c("id"))
names(m2)<-c("id","ti","value")
m2$cluster=2
m3<-melt(id3, id.vars=c("id"))
names(m3)<-c("id","ti","value")
m3$cluster=3
m4<-rbind(m1,m2,m3)
#Plotting
ggplot(data=m4,aes(x=factor(ti),y=value,color=id,group=id))+
  geom_point(size=2,aes(fill=id))+
  geom_line()+
  theme(axis.text.x = element_blank())+
  xlab("")+
  ylab("FPKM")+
  theme_bw()+  
  theme(axis.text.x = element_text(angle = 60,hjust=0,face = "bold", size = 12))+
  theme(axis.text.y = element_text(face = "bold", size = 12))+scale_x_discrete(position = "top")+
  theme(line = element_blank())+
  #using facet function to divid the plot into 3 subgroup
  facet_wrap(~cluster,nrow = 3)+
  #modify the figures' space
  theme(strip.text.x = element_blank())+theme(panel.spacing.y = unit(0,"lines"))+
  theme(legend.title=element_blank())

```
